Drought indices of water demand and supply, soil moisture, vegetation, and their impact on LULCC in continental Chile

### 2024-01-23 article/paper_drought_Chile2023.qmd
actualizar validacion era5

We validated the ERA5L variables for monthly mean temperature, total precipitation, and volumetric soil water content. For temperature and precipitation, we used the weather network from the Ministry of Agriculture of Chile (www.agromet.com). We used 266 stations located throughout Chile. For soil moisture, we select a soil network for water content and temperature, localized in the Aconcagua's river basin ("Norte Chico"). This network corresponds to 53 sensors of the type Temperature Moisture Sensor (TMS) [@Wild2019].

@@

The average metrics of performance of ERA5L over the 266 weather stations were in the case of temperature: $ubRMSE=1.06\,°C$, $MAE=1.131\,°C$, and $CC=0.963$. For precipitation, $MAE=28.1\,mm$, $Bias=1.93$, and $CC=0.845$.

>>>>

We compared the ERA5L variables for monthly mean temperature, total precipitation, and volumetric soil water content against values retrieved by weather stations. For temperature and precipitation, we used the weather network from the Ministry of Agriculture of Chile (www.agromet.com) between 2015 and 2023. We used 277 stations located throughout Chile. For soil moisture, we select a private soil network that is owned by the agricultural enterprise Garces Fruit, which has 99 stations in Central Chile, located in cherry fruit crops. We used daily data for the year 2022 and the months outside the growing season, May to September, to avoid the effect of irrigation on soil moisture, which is hardly captured by ERA5L.

@@

The average metrics of performance of ERA5L over the 266 weather stations were in the case of monthly temperature: $ubRMSE=1.06\,°C$, $MAE=1.131\,°C$, and $CC=0.963$, showing a good agreement, low error, and low overestimation. For cumulative monthly precipitation, $MAE=28.1\,mm$, $bias=1.93$, and $CC=0.845$, showing a high correlation and a 93% bias and being overestimated by ERA5L. In the case of the 97 soil moisture stations, we averaged for the three depths (30, 60, and 90m) and then compared it with volumetric water content at 1m derived from ERA5L. For this case, we made a daily comparison, having a $CC=0.71$, $RMSE=0.174\,m^3m^{-3}$, $MAE=0.167\,m^3m^{-3}$, and $bias=1.74$. The ERA5 soil moisture overestimate is 74%, but it has a kind of good correlation.

### 2024-01-23 article/paper_drought_Chile2023.tex
actualizar validacion era5

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable]}{\end{tcolorbox}}\fi

@@

\hypertarget{in-situ-data}{%
\subsubsection{in-situ data}\label{in-situ-data}}

POR COMPLETAR

@@

POR COMPLETAR

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

\hypertarget{weather-stations}{%
\subsubsection{Weather stations}\label{weather-stations}}

We compared the ERA5L variables for monthly mean temperature, total
precipitation, and volumetric soil water content against values
retrieved by weather stations. For temperature and precipitation, we
used the weather network from the Ministry of Agriculture of Chile
(www.agromet.com) between 2015 and 2023. We used 277 stations located
throughout Chile. For soil moisture, we select a private soil network
that is owned by the agricultural enterprise Garces Fruit, which has 99
stations in Central Chile, located in cherry fruit crops. We used daily
data for the year 2022 and the months outside the growing season, May to
September, to avoid the effect of irrigation on soil moisture, which is
hardly captured by ERA5L.

@@

To account for the performance of the ERA5L climatic variables regarding
the values measured by the weather stations. We selected the following
metrics:

\[MAE = \frac{1}{n}\sum |{E-S}|\] \[Bias = \frac{\sum E}{\sum S}\]
\[ubRMSE =\sqrt{\frac{\sum{ \left[ (E_i-\overline{E})-(S_i-\overline{S}) \right ] ^2}}{n}}\]

\[CC = \frac{\sum (S_i-\bar{S})(E_i-\bar{E})}{\sqrt{(Si-\bar{S})^2(E_i-\bar{E})^2}}\]

\(MAE\): mean absolute error \(bias\): bias \(ubRMSE\): unbiassed root
mean squared error \(CC\): coefficient of correlation \(S\): value of
the variable measure by the weather station \(E\): value of the variable
measure by ERA5L

@@

The average metrics of performance of ERA5L over the 266 weather
stations were in the case of monthly temperature: \(ubRMSE=1.06\,°C\),
\(MAE=1.131\,°C\), and \(CC=0.963\), showing a good agreement, low
error, and low overestimation. For cumulative monthly precipitation,
\(MAE=28.1\,mm\), \(bias=1.93\), and \(CC=0.845\), showing a high
correlation and a 93\% bias and being overestimated by ERA5L. In the
case of the 97 soil moisture stations, we averaged for the three depths
(30, 60, and 90m) and then compared it with volumetric water content at
1m derived from ERA5L. For this case, we made a daily comparison, having
a \(CC=0.71\), \(RMSE=0.174\,m^3m^{-3}\), \(MAE=0.167\,m^3m^{-3}\), and
\(bias=1.74\). The ERA5 soil moisture overestimate is 74\%, but it has a
kind of good correlation.
